SMT-Boosted Security Types for Low-Level MPC

### 2025-11-17 toplas26/model.tex
checkpoint

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X_1$ given $X_2$ where $X_1 \cup X_2 \subseteq \dom(\pmf)$ and $X_1 \cap X_2 = \varnothing$
  is denoted $\condd{\pmf}{X_1}{X_2}$ and defined as follows:
  $$
  \forall \store \in \mems(X_1 \cup X_2)\ .\ 
  \condd{\pmf}{X_1}{X_2}(\store) \defeq
  \begin{cases}
    \begin{array}{ll}
      0 & \text{\ if\ } \margd{\pmf}{X_2}(\store_{X_2}) = 0\\
      \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ & \text{\ o.w.}
     \end{array}
  \end{cases}
  $$
\end{definition}
We also define some convenient syntactic sugarings. The first will allow us to
compare marginal distributions under different realization conditions
(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
\begin{definition}
  Given $\pmf$, for all $\store_1 \in \mems(X_1)$ and $\store_2 \in \mems(X_2)$ define:
  \begin{enumerate}
  \item $\condd{\pmf}{X_1}{\store_2}(\store_1) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \end{enumerate}
\end{definition}
As in previous work \cite{darais2019language,barthe2019probabilistic,skalka-near-ppdp24,li2023lilac}, 
probabilistic independence, aka separation, is an important concept and we adopt the
following standard notation to express conditional and unconditional separation:
\begin{definition}
%  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
%  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
    $\store \in \mems(X_1 \cup X_2)$ and $\store' \in \mems(X_3)$ we have
  $\pmf(\store|\store') =
  \pmf(\store_{X_1}|\store') * \pmf(\store_{X_2}|\store')$. If $X_1$ is empty
  we write $\sep{\pmf}{X_2}{X_3}$.
\end{definition}

@@

We treat all elements of $\runs(\prog)$ as equally likely.  This
establishes the basic program distribution that can be marginalized
and conditioned to quantify input/output information dependencies.
%In this
%setting, given a program $\prog$ with $\iov(\prog) = S \cup V$ and
%$\flips(\prog) = F$ we will consider all $\store \in \mems(S \cup V
%\cup F)$ such that $ \config{\store_{S \cup F}}{\prog} \redxs
%\config{\store_}{\varnothing} $ to be equally probable, establishing
%the basic distribution of the program. %From this, we can immediately
%derive the marginal distribution of $S \cup V$ to reason about
%dependencies between secrets and views.
\begin{definition}
  \label{def-progtt}
  \label{def-progd}
  \label{definition-progd}
  The \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
  defined such that for all $\store \in \mems(\vars(\prog))$:
  $$
  \progtt(\prog)(\store) =  1 / |\runs(\prog)| \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
  $$
  For any $\adversary$ the basic \emph{$\adversary$ distribution of $\prog$}, written
  $\progtt(\prog,\adversary)$, is
  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
  $$
  \progtt(\prog,\adversary)(\store) =  1 / |\botruns(\prog)| \ \text{if}\ \store \in \botruns(\prog), \text{otherwise}\ 0
  $$
  where $\botruns$ pads out undefined views and outputs with $\bot$:
  $$
  \begin{array}{c}
    \botruns(\prog) \defeq \\
    \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid 
    \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
  \end{array}
  $$
\end{definition}

@@

Information about honest secrets can be revealed to corrupt clients
through messages sent from honest to corrupt clients, and through
publicly broadcast information from honest clients. Dually,
corrupt clients can impact protocol integrity through the messages
sent from corrupt to honest clients, and through publicly broadcast information
from corrupt clients. We call the former \emph{corrupt views}, and
the latter \emph{honest views}. We let $V$ range over sets
of views, i.e., subsets of $M \cup P$.
\begin{definition}[Corrupt and Honest Views]
  We let $V$ range over \emph{views} which are sets of messages
  and reveals. Given a program $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$,
  define $\views(\prog) \defeq M \cup P$, and define $\houtputs$ as
  the messages and reveals in $V = M \cup P$ sent from honest to corrupt
  parties, called \emph{corrupt views}:
  $$
  \begin{array}{lcl}
    \houtputs & \defeq
        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in H \ \}\ \cup \\
      & & \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
           \prog \wedge \cid \in C \wedge \cid' \in H \ \} 
  \end{array}
  $$
  and similarly define $\cinputs$ as the subset of $V$ sent from corrupt to honest
  parties, called \emph{honest views}:
  $$
  \begin{array}{lcl}
    \cinputs &  \defeq
        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in C \ \} \ \cup\\
      & & \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
              \prog \wedge \cid \in H \wedge \cid' \in C \ \}
  \end{array}
  $$
\end{definition}

@@

Since MPC protocols release some information about secrets through
outputs of $\idealf$, they do not enjoy strict noninterference.
Public reveals and protocol outputs are fundamentally forms of
declassification. But the following property of probabilistic
noninterference \emph{conditioned on output} is sound for passive
security \cite{skalka-near-ppdp24}.
\begin{definition}[Noninterference modulo output]
  \label{definition-NIMO}
  We say that a program $\prog$ with $\iov(\prog) = S \cup V \cup O$
  satisfies \emph{noninterference modulo output}
  iff for all $H$ and $C$ and $\store_1 \in \mems(S_C \cup O)$ and $\store_2 \in \mems(\houtputs)$
  we have:
  $$
  \condd{\progtt(\prog)}{S_H}{\store_1} = \condd{\progtt(\prog)}{S_H}{\store_1 \uplus \store_2}
 $$
\end{definition}
This conditional noninterference property implies that
corrupt views give the adversary no better chance of guessing honest
secrets than just the output and corrupt inputs do.

@@

\emph{Integrity} is an important hyperproperty in security models that
admit malicious adversaries. Consistent with formulations in
deterministic settings, we define protocol integrity as the
preservation of high equivalence (of secrets and views). Intuitively,
this property says that any adversarial strategy either ``mimics'' a
passive strategy with some choice of inputs or causes an abort.
\begin{definition}[Integrity]
  \label{def-integrity}
  We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
  \emph{integrity} iff for all $H$, $C$, and $\adversary$,
  if $\store \in \aruns(\prog)$ 
  then there exists $\store' \in \mems(S)$ with $\store_{S_H} = \store'_{S_H} $ and:
    $$
    \condd{\progtt(\prog,\adversary)}{X}{\store_{S_H \cup \cinputs}} =
    \condd{\progtt(\prog)}{X}{\store'}
    $$ 
  where $X \defeq (\houtputs \cup O_H) \cap \dom(\store)$. 
\end{definition}
Integrity plus noninterference modulo output implies malicious security
in the real/ideal model \cite{skalka-near-ppdp24}, and integrity plus
gradual release is a probabilistic form of robust declassification
\cite{sabelfeld2009declassification}.

>>>>

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X_1$ given $X_2$ where $X_1 \cup X_2 = \dom(\pmf)$ 
  is denoted $\condd{\pmf}{X_1}{X_2}$ and defined as follows:
  $$
  \forall \store \in \mems(X_1 \cup X_2)\ .\ 
  \condd{\pmf}{X_1}{X_2}(\store) \defeq
  \begin{cases}
    \begin{array}{ll}
      0 & \text{\ if\ } \margd{\pmf}{X_2}(\store_{X_2}) = 0\\
      \pmf(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ & \text{\ o.w.}
     \end{array}
  \end{cases}
  $$
\end{definition}
We also define some convenient syntactic sugaring for ``implicit
marginalization'' given a realization.
%The first will allow us to
%compare marginal distributions under different realization conditions
%(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
\begin{definition}
  Given $\pmf$ and $X \subseteq \dom(\pmf)$, for all $\store \in \mems(X)$
  define $\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$.
  %and
  %$\store_2 \in \mems(X_2)$ define
  %$\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
  %\begin{enumerate}
  %\item $\condd{\pmf}{X_1}{\store_2}(\store_1) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  %\item $\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
  %\item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  %\end{enumerate}
\end{definition}
As in previous work \cite{darais2019language,barthe2019probabilistic,skalka-near-ppdp24,li2023lilac}, 
probabilistic independence, aka separation, is an important concept and we adopt the
following standard notation to express conditional and unconditional separation:
\begin{definition}
%  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
%  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
    $\store \in \mems(X_1 \cup X_2)$ and $\store' \in \mems(X_3)$ we have
  $\pmf(\store|\store') =
  \pmf(\store_{X_1}|\store') * \pmf(\store_{X_2}|\store')$. If $X_1$ is empty
  we write $\sep{\pmf}{X_2}{X_3}$.
\end{definition}

@@

We treat all elements of $\runs(\prog)$ as equally likely.  This
establishes the basic program distribution that can be marginalized
and conditioned to quantify input/output information dependencies.
%In this
%setting, given a program $\prog$ with $\iov(\prog) = S \cup V$ and
%$\flips(\prog) = F$ we will consider all $\store \in \mems(S \cup V
%\cup F)$ such that $ \config{\store_{S \cup F}}{\prog} \redxs
%\config{\store_}{\varnothing} $ to be equally probable, establishing
%the basic distribution of the program. %From this, we can immediately
%derive the marginal distribution of $S \cup V$ to reason about
%dependencies between secrets and views.
\begin{definition}
  \label{def-progtt}
  \label{def-progd}
  \label{definition-progd}
  The \emph{basic distribution of $\prog$}, written $\progtt(\prog)$, is
  defined such that for all $\store \in \mems(\vars(\prog))$:
  $$
  \progtt(\prog)(\store) =  1 / |\runs(\prog)| \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
  $$
  For any $\adversary$ the basic \emph{$\adversary$ distribution of $\prog$}, written
  $\progtt(\prog,\adversary)$, is
  defined such that for all $\store \in \mems(\iov(\prog) \cup R)$:
  $$
  \progtt(\prog,\adversary)(\store) =  1 / |\botruns(\prog)| \ \text{if}\ \store \in \botruns(\prog), \text{otherwise}\ 0
  $$
  where $\botruns$ pads out undefined views and outputs with $\bot$:
  $$
  \begin{array}{c}
    \botruns(\prog)\ \defeq\ % \\
    \{ \store\{ x_1 \mapsto \bot, \ldots, x_n \mapsto \bot \} \mid 
    \store \in \aruns(\prog) \wedge \{ x_1,\ldots,x_n\} = (V \cup O) - \dom(\store) \}
  \end{array}
  $$
\end{definition}

@@

Information about honest secrets can be revealed to corrupt clients
through messages sent from honest to corrupt clients, and through
publicly broadcast information from honest clients. Dually,
corrupt clients can impact protocol integrity through the messages
sent from corrupt to honest clients, and through publicly broadcast information
from corrupt clients. We call the former \emph{corrupt views}, and
the latter \emph{honest views}. We let $V$ range over sets
of views, i.e., subsets of $M \cup P$.
\begin{definition}[Corrupt and Honest Views]
  We let $V$ range over \emph{views} which are sets of messages
  and reveals. Given a program $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$,
  define $\views(\prog) \defeq M \cup P$, and define $\houtputs$ as
  the messages and reveals in $V = M \cup P$ sent from honest to corrupt
  parties, called \emph{corrupt views}:
  $$
  \begin{array}{lcl}
    \houtputs & \defeq
        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in H \ \}\ \cup\ %\\
    % & &
    \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
           \prog \wedge \cid \in C \wedge \cid' \in H \ \} 
  \end{array}
  $$
  and similarly define $\cinputs$ as the subset of $V$ sent from corrupt to honest
  parties, called \emph{honest views}:
  $$
  \begin{array}{lcl}
    \cinputs &  \defeq
        & \{\ \rvl{w} \mid\ \reveal{w}{\be}{\cid} \in \prog \wedge \cid \in C \ \} \ \cup\ %\\
    %& &
    \{\ \elab{\mesg{w}}{\cid}\ \mid\  \eassign{\mesg{w}}{\cid}{\be}{\cid'} \in
              \prog \wedge \cid \in H \wedge \cid' \in C \ \}
  \end{array}
  $$
\end{definition}

@@

Since MPC protocols release some information about secrets through
outputs of $\idealf$, they do not enjoy strict noninterference.
Public reveals and protocol outputs are fundamentally forms of
declassification. But the following property of probabilistic
noninterference \emph{conditioned on output} is sound for passive
security \cite{skalka-near-ppdp24}.
\begin{definition}[Noninterference modulo output]
  \label{definition-NIMO}
  We say that a program $\prog$ with $\iov(\prog) = S \cup V \cup O$
  satisfies \emph{noninterference modulo output}
  iff for all $H$ and $C$ %and $\store_1 \in \mems(S_C \cup O)$ and $\store_2 \in \mems(\houtputs)$
  we have:
  $$
  \condd{\progtt(\prog)}{S_H}{S_C \cup O} = \condd{\progtt(\prog)}{S_H}{S_C \cup O \cup \houtputs}
 $$
\end{definition}
This conditional noninterference property implies that
corrupt views give the adversary no better chance of guessing honest
secrets than just the output and corrupt inputs do.

@@

\emph{Integrity} is an important hyperproperty in security models that
admit malicious adversaries. Consistent with formulations in
deterministic settings, we define protocol integrity as the
preservation of high equivalence (of secrets and views). Intuitively,
this property says that any adversarial strategy either ``mimics'' a
passive strategy with some choice of inputs or causes an abort.
\begin{definition}[Integrity]
  \label{def-integrity}
  We say that a protocol $\prog$ with $\iov(\prog) = S \cup V \cup O$ has
  \emph{integrity} iff for all $H$, $C$, and $\adversary$,
  if $\store \in \aruns(\prog)$ we have:
    $$
    \condd{\progtt(\prog,\adversary)}{X}{S_H \cup \cinputs} =
    \condd{\progtt(\prog)}{X}{S_H \cup \cinputs}
    $$ 
  where $X \defeq (\houtputs \cup O_H) \cap \dom(\store)$. 
\end{definition}
Integrity plus noninterference modulo output implies malicious security
in the real/ideal model \cite{skalka-near-ppdp24}, and integrity plus
gradual release is a probabilistic form of robust declassification
\cite{sabelfeld2009declassification}.

### 2025-11-17 toplas26/sty.tex
checkpoint

@@

\begin{lemma}
  \label{lemma-sty-noninteractive-sound}
  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
    \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
    \end{enumerate}
\end{lemma}

>>>>

$$
\mems(\Phi) \defeq \{ \store \mid \dom(\store) = \setit{\ x_{\phi}\ \mid\ \phi \in \Phi\ } \}
$$

\begin{definition}
  Given pmf $\pmf$ and $\Phi$ with $\dom(\pmf) = \bigcup_{\phi \in \Phi} \dom(\phi)$,
  for all $\store \in \mems(\Phi)$ define:
%$$
%\pmf(\phi = v) \defeq \sum_{\store \in \mems(X) \ \text{with}\ \store(\phi) = v} \pmf(\store)
%$$
  $$
  \pmf(\store) \defeq \sum_{\store' \in \mems(X)}
  \begin{cases}
    \pmf(\store') \ \text{if}\ \forall \phi \in \Phi . \store'(\phi) = \store(x_\phi)  \\
    0 \ \text{otherwise}
  \end{cases}
  $$
\end{definition}

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X$ given $\Phi$ where $\bigcup_{\phi \in \Phi} \dom(\phi) \cup X = \dom(\pmf)$ 
  is denoted $\condd{\pmf}{X}{\Phi}$ and defined as follows: 
  $$
  \forall \store_1 \in \mems(X)\ .\ \forall \store_2 \in \mems(\Phi) \ .\ 
  \condd{\pmf}{X}{\Phi}(\store_1 \uplus \store_2) \defeq
  \begin{cases}
    0 \text{\ if\ } \margd{\pmf}{\Phi}(\store_2) = 0\\
    \pmf(\store_1 \uplus \store_2) / \pmf(\store_2) \text{\ otherwise}
  \end{cases}
  $$
\end{definition}

@@

\begin{lemma}
  \label{lemma-sty-noninteractive-sound}
  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
    \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\phi_i^1 + \phi_i^2, \phi_i^a}}$
    \end{enumerate}
\end{lemma}
